---
title: "Benchmarking immune signatures honestly, in gastric cancer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking immune signatures honestly, in gastric cancer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A signature that separates outcomes is not thereby informative: randomly chosen genes
of the same size often do the same. `immunoTriage` makes the two comparisons that decide
the question — against a matched random null, and against what any model could achieve
on the same data.

```{r setup}
library(immunoTriage)
data(gc_ici); data(gc_tcga); data(gastro_signatures); data(gastro_controls)
sigs <- c(gastro_signatures, gastro_controls)
```

## Prognosis, in untreated gastric cancer

```{r prognosis}
y <- survival::Surv(gc_tcga$time, gc_tcga$event)
prog <- benchmark_signatures(gc_tcga$expr, y, sigs, n_random = 200, seed = 1)
prog[order(-prog$delta), c("signature", "performance", "random_median", "delta", "beats_null")]
```

Note where the confounder controls land. If a proliferation signature outranks the
immune signatures, the prognostic signal is not immune-specific.

## Response, in checkpoint-treated gastric cancer

```{r response}
resp <- benchmark_signatures(gc_ici$expr, gc_ici$response, sigs, n_random = 200, seed = 1)
resp[order(-resp$delta), c("signature", "performance", "random_median", "delta", "beats_null")]
```

## The dissociation

```{r dissoc}
d <- test_dissociation(prog, resp, gastro_families,
                       target = "immune", confounder = "confounder")
d$per_family
d$interaction_p
```

The immune family shifting between settings is only interpretable if the confounder
family does not. `test_dissociation()` reports both so the comparison cannot be made
by accident.

## What is the ceiling?

```{r ceiling, eval=FALSE}
x  <- t(gc_ici$expr)
cl <- estimate_ceiling(x, gc_ici$response, models = c("enet", "rf"), nfolds = 5)
cl$per_model; cl$ceiling
compare_to_ceiling(resp, cl)
```

If the ceiling sits close to the best signature, the task is information-limited and
a more complex model will not help. Reporting that is more useful than publishing
another score that differs by a rounding error.
